[Re] A neural model of the saccade generator in the reticular formation

### 2018-02-21 article/senden-schuecker-hahne-diesmann-goebel-2018.tex
updates according to Tiziano Zito's suggestions

We provide an implementation of the saccade generator (SG); a rate
neuron model of neural circuitry in the reticular formation proposed by
Gancarz \& Grossberg \autocite{Gancarz1998}. The SG is an integral part
of the eye movement system \autocite{Grossberg2012} and as such vital
for developing large-scale architectures of visuo-motor integration. For
this reason it is of interest to implement the model in a publicly
available, widely used, and actively developed neural simulation
framework such as NEST \autocite{Gewaltig2007}. We show that the model
translates well to the NEST framework as our implementation faithfully
reproduces all simulation results reported in the original publication.
Our code uses the Python interface \autocite{Eppler2008} for legibility
with both model and analysis scripts being implemented using Python
2.7.12.

@@

\begin{equation}
\begin{array}{ll}
\tau\frac{dE_l}{dt} &= -3.5E_l+(2-E_l)(5E_l+1)-(E_l+1)(10L_r+20g(P))) \\\\
\tau\frac{dE_r}{dt} &= -3.5E_r+(2-E_r)(5E_r+1)-(E_r+1)(10L_l+20g(P))) \\
\end{array}
\label{eq:ebn}\end{equation} with a nonlinear gain function \(g(\cdot)\)
given by

@@

In all simulations we used the Exponential Euler method for numerical
integration of rate neurons \autocite{Hahne2017} at a time step of
\(0.05\,\mathrm{ms}\) and a time constant of \(50\,\mathrm{ms}\). It
should be noted that rates of all neurons were initialized to zero and
were thus not at resting equilibrium. In order to address this, the
model was allowed to evolve for \(100\,\mathrm{ms}\), to relax towards
equilibrium before applying any input. Furthermore, we always simulated
the full model; i.e.~both its horizontal and vertical components even if
input was applied only to one of the two.

@@

In conclusion, our reproduction confirms the results of the original
publication and shows that an implementation in the NEST framework is
feasible. This allows for the straightforward integration of the saccade
generator into models of the visuo-motor system forming the interface
between sensory processing and motor control.

@@

>>>>

In 2013 the European Comission launched the Human Brain Project
(\href{https://www.humanbrainproject.eu/en/}{HBP}) tasked to build a
research infrastructure spurring pan-european collaboration.
Collaboration at such a scale, involving both empirically-oriented and
theoretical neuroscientists, offers the unique opportunity to develop
large-scale models of the brain. Specifically, individual research
groups might develop models of distinct brain regions which can
subsequently be combined into a unified whole. To facilitate this
colaboration, the HBP encourages the utilization of publicly available,
widely used, and actively developed neural simulation frameworks such as
NEST \autocite{Gewaltig2007}. In light of this, the NEST framework has
recently been extended to support simulation of functionally inspired
rate neuron models in addition to biologically grounded spiking neuron
models \autocite{Hahne2017}. Here we make use of this new functionality
of the NEST framework (v2.16.0) to provide an implementation of the
saccade generator (SG); a rate neuron model of neural circuitry in the
reticular formation proposed by Gancarz \& Grossberg
\autocite{Gancarz1998}. The SG is an integral part of the eye movement
system \autocite{Grossberg2012} and as such vital for developing
large-scale architectures of visuo-motor integration. We show that the
model translates well to the NEST framework as our implementation
faithfully reproduces all simulation results reported in the original
publication. Our code uses the Python interface \autocite{Eppler2008}
for legibility with both model and analysis scripts being implemented
using Python 2.7.12.

@@

\begin{equation}
\begin{array}{ll}
\tau\frac{dE_l}{dt} &= -3.5E_l+(2-E_l)(5L_l+1)-(E_l+1)(10L_r+20g(P))) \\\\
\tau\frac{dE_r}{dt} &= -3.5E_r+(2-E_r)(5L_r+1)-(E_r+1)(10L_l+20g(P))) \\
\end{array}
\label{eq:ebn}\end{equation} with a nonlinear gain function \(g(\cdot)\)
given by

@@

In all simulations we used the Exponential Euler (EE) method for
numerical integration of all except of tonic neurons
\autocite{Hahne2017} at a time step of \(0.05\,\mathrm{ms}\) and a time
constant of \(50\,\mathrm{ms}\). Tonic neurons were integrated using the
Euler-Maruyama (EM) method since they do not exhibit passive decay
leading to division-by-zero when using the EE method. It should be noted
that initial firing rates of neurons were not at resting equilibrium. In
order to address this, the model was allowed to evolve for
\(100\,\mathrm{ms}\), to relax towards equilibrium before applying any
input. Furthermore, as in the original publication, all activations were
bounded from below at zero. Finally, we always simulated the full model;
i.e.~both its horizontal and vertical components even if input was
applied only to one of the two.

@@

In conclusion, our reproduction confirms the results of the original
publication and shows that an implementation in the NEST framework is
feasible. This allows for the straightforward integration of the saccade
generator with computational models of other components of the
visuo-motor system (e.g.~salience computation) within a shared
framework.

@@

\section{Appendix}\label{appendix}

\subsection{A1 - Rate neurons in NEST}\label{a1---rate-neurons-in-nest}

For a detailed, comprehensive, treatment of rate-based neuron models in
NEST see Hahne et al.\autocite{Hahne2017}. Here we provide only a brief
overview to facilitate readability/interpretability of the code provided
with this publication. In general, rate-based neurons in NEST consist of
two components; a neuron and a gain function.

\subsubsection{Neuron}\label{neuron}

Two types of neuron base classes exist in NEST v2.16.0:
\texttt{rate\_neuron} and\\
\texttt{rate\_transformer\_node}. The \texttt{rate\_neuron} implements
continuous rate dynamics with and without multiplicative coupling. It
can apply a nonlinear gain function to its inputs either before or after
their weighted summation. It is numerically integrated using the EE
method unless a neuron does not exhibit exponential decay in which case
the EM method is used. Furthermore, rates can be bounded (from below at
zero) or unbounded. Finally, noise (in the form of a Wiener process) can
be added to its input or to its output. The \texttt{rate\_neuron} has a
total of seven parameters (see table \ref{tbl:table_2}) and two further
properties \texttt{rate} (current firing rate) and \texttt{noise}
(current noise level).

\begin{longtable}[]{@{}lll@{}}
\caption{\texttt{rate\_neuron} parameters.
\label{tbl:table_2}}\tabularnewline
\toprule
parameter & default value & description\tabularnewline
\midrule
\endfirsthead
\toprule
parameter & default value & description\tabularnewline
\midrule
\endhead
\texttt{tau} & \(10\,\mathrm{ms}\) & time constant\tabularnewline
\texttt{lambda} & \(1\) & passive decay rate\tabularnewline
\texttt{std} & \(1\) & noise standard deviation\tabularnewline
\texttt{mean} & \(0\) & mean firing rate\tabularnewline
\texttt{linear\_summation} & \texttt{true} & if true, apply gain
function after summation\tabularnewline
\texttt{rectify\_output} & \texttt{false} & if true, multiplicative
coupling\tabularnewline
\texttt{mult\_coupling} & \texttt{false} & if true, rate is
bounded\tabularnewline
\bottomrule
\end{longtable}

For the implementation of this neuron see
\texttt{rate\_neuron\_ipn\_impl.h} and\\
\texttt{rate\_neuron\_opn\_impl.h} in
\texttt{../nest/nest-simulator-master/models} for input and output
noise, respectively.

The \texttt{rate\_transformer\_node} instantaneously applies a
nonlinearity to its inputs but does not exhibit temporal dynamics. It
can be used to apply different nonlinearities to different inputs to a
single neuron. This class has only the \texttt{rate} property and no
parameters. For its implementation see
\texttt{rate\_transformer\_node\_impl.h} in
\texttt{../nest/nest-simulator-master/models}.

\subsubsection{Gain function}\label{gain-function}

Five gain functions exist in NEST v2.16.0: \texttt{lin\_rate},
\texttt{sigmoid\_rate},\\
\texttt{sigmoid\_rate\_gg\_1998}, \texttt{tanh\_rate}, and
\texttt{threshold\_lin\_rate}. Documentation of each gain function is
given in the header file with the corresponding name within
\texttt{../nest/nest-simulator-master/models}. We added
\texttt{sigmoid\_rate\_gg\_1998} specifically for the reimplementation
of the saccade generator presented here.

Gain functions are combined with a neuron class to obtain a specific
neuron model. For instance, the combination of the \texttt{lin\_rate}
gain function with the \texttt{rate\_neuron\_ipn} class gives the
\texttt{lin\_rate\_ipn} neuron model. The parameters of a model are a
combination of all parameters of the neuron base class and all
parameters of the gain function. Within Python the command
\texttt{nest.Models()} lists all existing model types including all
implemented combinations of gain function and neuron class.
